MLPPI Wizard: An Automated Multi-level Partitioning Tool on Analytical Workloads

### 2017-03-16 mlppi-comsis17/paper.tex
Experiment Setup rewritten

In this section we analyze the time complexity of 
the proposed preprocessing, initial, and optimized phases.
For the time complexity we compute the logical running time of the wizard, using the {\em number of API calls} 
made to the server during the phases.

@@

{\bf Proof}. Let us consider a worst case such that given a workload of 
$N$ queries, each query references all the fields of the fact table, and 
each field is associated with at most $V$ values by the predicates of the queries. 
Each query may have a single-value range. 
A range consisting of only a single value per field can be mapped to each query. 
Thus, the running time complexity for the map construction 
is $\BigO$($N${$\cdot$}$C${$\cdot$}$V$). \qed

But we empirically observed that the bound of $\BigO$($N${$\cdot$}$C${$\cdot$}$V$) 
was overly pessimistic.

@@

\begin{lemma}
\label{lemma:init_phase}  
The running time complexity for the initial or optimized phases is 
$\BigO$($M^{2}${$\cdot$}$N$), 
where $M$ is the number of range pairs, and $N$ is the number of queries in a workload. 
\end{lemma}

{\bf Proof}. Suppose that the merge of every range pair influences all the queries. 
Every iteration, either the scan or query costs need to be recomputed 
for each query. 
In the first round the \hbox{re-computation} cost is paid as many times as $M${$\cdot$}$N$, 
and a chosen range pair is merged. In the subsequent round 
$M$-1 range pairs remains, and the following cost amounts to ($M$-1){$\cdot$}$N$. 
In the worst case we may end up consolidating all range pairs, thus having no partition on 
a target table. The total calls to the server can be increased up to

@@

Again, our experiments showed that $\BigO$($M^{2}${$\cdot$}$N$), the running time complexity, 
was overly pessimistic. Specifically, 
the number of calls made in our experiments was much fewer than that bound, as 
it was almost rare for a range pair to involve all queries in a workload 
(as seen in Table~\ref{tab:workload_stat}).

In the following, we evaluate the proposed algorithm.

@@

In this section, we describe our environment settings, 
report the statistics measured in our experiments, and 
demonstrate the performance of the MLPPI wizard (WIZARD), compared with 
that of no partitioning and the partitioning by a human expert (EXP).

@@

The MLPPI wizard was implemented as a prototype 
on top of the Teradata DBMS server. 
It was written in Java. 
The performance of the wizard was evaluated 
on the Teradata DBMS server machine running Unix.

When it comes to workload generation, we took advantage of 
our simple star schema query \hbox{generator}. 
The generator assumes that 1) available operators, 2) fields, 
and 3) the minimum and \hbox{maximum} values of the fields are already known. 
For each query, the generator first randomly selects the number of single-table 
predicates to create. 
Then, it arbitrarily chooses a field and a specific operator 
for each predicate. 
If {\tt IN} operator on a chosen field is picked up, the query generator 
determines the number of values to add to the {\tt IN} list and then chooses random values within the min-max value 
range of the field. 
In this way the generator builds a query involving the generated predicates.

A generated query is based on a template that joins {\tt LINEORDER} and {\tt DDATE} with constraints defined by the predicates. 
This template is common in customer cases like reports and form 
templates. 
Using the query generator we generated two workloads consisting of 10 queries (10Q) and 20 queries (20Q).

For the query workloads we populated the fact table with a scale of 1TByte (1TB) and 3TBytes (3TB).

>>>>

In this section we analyze the time complexity of 
the proposed algorithms: preprocessing, initial, and optimized phases. 
The metric of this complexity analysis on each phase is the logical running time of the wizard, 
equivalent to the number of API calls made to the database server during that phase.

@@

{\bf Proof}. Consider the worst case such that given a workload of 
$N$ queries, each query references all the fields of the fact table, and 
each field is associated with at most $V$ values by the predicates of the queries. 
Each query may have a single-value range. 
A range consisting of only one value per field can be mapped to each query. 
Thus, the running time complexity for the map construction 
is $\BigO$($N${$\cdot$}$C${$\cdot$}$V$). \qed

But note that in our experiments the bound of $\BigO$($N${$\cdot$}$C${$\cdot$}$V$) 
was overly pessimistic.

@@

\begin{lemma}
\label{lemma:init_phase}  
The running time complexity for the initial (or optimized) phase is 
$\BigO$($M^{2}${$\cdot$}$N$), 
where $M$ is the number of range pairs, and $N$ is the number of queries in a workload. 
\end{lemma}

{\bf Proof}. Suppose that the merge of every range pair influences all the queries. 
Every iteration, either the scan or query costs need to be recomputed for each query. 
In the first round we pay the \hbox{re-computation} cost of $M${$\cdot$}$N$ 
and merge a victim range pair. In the subsequent round, $M$-1 range pairs remain, 
so the cost amounts to ($M$-1){$\cdot$}$N$. 
In the worst case we end up consolidating all range pairs, thus having no partition on 
the target table. The total calls made to the server can be added up to

@@

Again, our experiments showed that $\BigO$($M^{2}${$\cdot$}$N$), the running time complexity, 
was overly pessimistic. Specifically, 
the number of calls made in our experiments was by far fewer than that bound. 
That was because it was very rare for a range pair to involve all queries in a given workload 
(as seen in Table~\ref{tab:workload_stat}).

%In the following we evaluate the performance of our wizard using the proposed algorithms.

@@

In this section we evaluate the performance of our wizard using the proposed algorithms.
We first describe our environment settings. 
We then report the statistics measured during our experiments. 
Finally, we compare the performance of the recommendation 
by the MLPPI wizard with (1)~that of no partitioning 
and (2)~the partitioning of a human expert (EXP) under different setups.

@@

\paragraph{Development:} 
We implemented our MLPPI wizard as a prototype on top of the \hbox{Teradata} DBMS server. 
The wizard was written in Java. 
We evaluated the performance of the \hbox{wizard} 
on the Teradata DBMS server machine running Unix.

\paragraph{Workload:} When it comes to workload generation, we took advantage of 
our simple star schema query \hbox{generator}. 
The generator assumes that (i)~available operators, (ii)~fields, 
and (iii)~the minimum and \hbox{maximum} values of the fields are already known. 
For each query, the generator first randomly selects the number of single-table 
predicates to create. 
The generator then arbitrarily chooses a field and a specific operator 
for each predicate. 
If {\tt IN} operator on a chosen field is picked up, the query generator 
determines the number of values to add to the {\tt IN} list and then chooses random values within the min-max value 
range of the field. 
In this way the generator makes a query involving the generated predicates.

A generated query is based on a template that joins {\tt LINEORDER} and 
{\tt DDATE} with constraints defined by the predicates. 
This template is common in customer cases like reports and form 
templates. 
Using the query generator we generated two workloads consisting of 10 queries~(10Q) and 20 queries~(20Q).

For the query workloads we populated the fact table with a scale factor of 1TByte~(1TB) and 3TBytes~(3TB).
